Deriving Multi-Headed Projective Dependency Parses \\ from Link Grammar Parses

### 2014-05-27 doc/LinksToDAG.tex
started editing paper; added refs to bibfile

\newcommand{\Note}[1]{}
\renewcommand{\Note}[1]{\hl{[#1]}}  % comment out this definition to suppress all Notes                                                                                
\newcommand{\TODO}[1]{\Note{TODO: #1}}
\newcommand{\NoteSigned}[3]{{\sethlcolor{#2}\Note{#1: #3}}}
\newcommand{\NoteJE}[1]{\NoteSigned{JE}{LightBlue}{#1}}
\newcommand{\NoteJH}[1]{\NoteSigned{JH}{YellowGreen}{#1}}

\title{Deriving Multi-headed Planar Dependency Parses from Link Grammar Parses}

\author{Juneki Hong and Jason Eisner\\
  Department of Computer Science \\
  Johns Hopkins University \\
  Baltimore, MD 21218, USA \\ 
  {\tt \{juneki,jason\}@cs.jhu.edu} \\
}

@@

\begin{abstract}
% Copied from the google doc
There has been recent theoretical work on multi-headed versions of dependency grammar(cite). Such formalisms can be more syntactically and semantically expressive. However, there aren't any corpora at present (check this) except for Buch-Kromann's work on Danish \cite{kromann2003danish}, so it is hard to investigate the benefit of such parsers or to work on making them faster or more accurate. To fill this gap, we observe that link grammar produces parses that are similar to multi-headed planar dependency parses except that the links are undirected. We use Integer Linear Programming to assign consistent directions to the links in a corpus of NNN parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Russian, and more. (what statistics does it use?). We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses. The resulting parses in English appear linguistically plausible, although they are not in general consistent with CoNLL-style parses of the same sentences; we discuss the differences. We also report ...
\end{abstract}

\section{Introduction}
\TODO{Intro}
Projective dependency parsing is the task of mapping a sentence to a directed acyclic tree. Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse. In this paper we explore whether these relationships include dependencies, and whether the multiheadedness of the link grammar offers additional dependency relationships not found in other corpora.

We will accomplish this by generating link parses using the AbiWord/CMU link grammar parser, learning a direction to every encountered link label type, and then using these assignments for all link tokens in the link parse corpus. We formulate an integer linear programming problem that will find the smallest set of assignments that will still produce dependency parses that are acyclic with all nodes reachable from the root.

From the ILP, we produce a dictionary of label-to-direction assignments. We publish this so that anyone can easily turn a link parse into a dependency parse.

Using the sentences of a dependency corpus as comparison, we find that the link parses do not wholly subsume dependency parses, and that the undirected links match roughly two thirds of the arcs in the conll dependency corpus. \TODO{This is because...} 
Of the links that match a corresponding dependency arc we measure whether the direction we chose is the same as the original dependency data. 
Finally, we look to see that given a link label, whether the original conll label can be recovered.

%about 20\%

%Kromann 2003 Danish dependency treebank\cite{kromann2003danish}
%McDonald and Pereira 2006\cite{McDonald2006}
%Sagae and Tsujii 2008\cite{Sagae2008}
%Gomez-Rodriguez and Nivre 2013\cite{gomez-rodriguez_divisible_2013}

@@

\section{Link Grammars}
Link grammars are a grammatical system equivalent to context-free grammars that assign linking requirements to every given word\cite{SleatorTemperly91}. A link parser then tries to satisfy all of these requirements for every word of a sentence while still maintaining planarity. The resulting links describe the relationships between constituents in a parse.

The link grammar is based on a set of handwritten dictionaries. Instead of going through these dictionaries, we learned the grammar using an ILP. This approach also potentially allows us to analyze link grammar dictionaries other than English. \TODO{explore other link grammar dictionaries}

\section{Integer Linear Programming}
We formulated an ILP to assign directionality to all of the links in a set of link grammar parses, while still mandating that the resulting dependency parses are connected DAGs, with the nodes reachable from the root.

In order to choose which direction the link labels are assigned, for every encountered label we generate two boolean label\_direction variables correspending to left and right. Setting the only one of these variables to \textsc{TRUE} chooses a consistent direction for a label throughout all parses, and because we assign a directionality to every link type at least one must be set to true. Setting both variables would allow us to go in either direction.

Our ILP objective is to minimize the number of label\_direction variables set to true, while still satisfying the acyclicity and reachability constraints. The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite (cite?).

\TODO{cite model minimization papers}

\subsection{Slack}
We introduced slack on the directional dependency variables such that the link types were allowed to deviate from the majority one percent of the time before the ILP would assign both directional dependency variables to true. This addressed noise in the link parser's label assignments, while still allowing for both directions to be assigned if needed.
\begin{figure}
%\centering
  \small
  \begin{align*}
    &\textsc{dir} = \{\textsc{left}, \textsc{right}\}, \textsc{l} \in \textsc{labels} & \\
    &\textsc{m} = \left|{\textsc{links}}\right|
  \end{align*}
  \begin{equation}
    \textsc{count}_{\textsc{l}, \textsc{dir}} = \sum \textsc{link}_{\textsc{l},\textsc{dir}}
  \end{equation}
  \begin{equation}
    \textsc{count}_{\textsc{l}, \textsc{dir}} = \textsc{m} \cdot \textsc{allowed}_{\textsc{l},\textsc{dir}} + \textsc{slack}_{\textsc{l}} 
  \end{equation}
  \begin{equation}
    \textsc{cost}_{\textsc{l}} = \frac{100}{\textsc{count}_{\textsc{l},\textsc{left}} + \textsc{count}_{\textsc{l},\textsc{right}}}
  \end{equation}
  \begin{equation}
    \text{minimize} \sum_{(\textsc{l}, \textsc{dir}} \textsc{allowed}_{\textsc{l},\textsc{dir}}) + \textsc{cost}_{\textsc{l}} \cdot \sum_\textsc{l} \textsc{slack}_{\textsc{l}} 
  \end{equation}
  \begin{align*}
    \text{subject to } &\forall_{i \in \textsc{sentence}}: \textsc{reachable}_{\textsc{root}, i} = \textsc{true} &\\
                       &\forall_{\substack{i,j \in \textsc{sentence} \\ i\neq j }}: \textsc{cycle}_{i,j} = \textsc{false} &
  \end{align*}

  \caption{\small The ILP formulation.}
  %\\L is a label\\M is the size of the corpus.}
\end{figure}

@@

\begin{figure}[ht!]
  \small
  \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
  \caption{\small Precision and recall of increasing runs of English sentences}
\end{figure}

@@

%\section{Other languages}

\section{Link Corpus Experiment and Results}
We ran the link parser on the english bnews corpus sentences. \footnote{We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph. This happened for roughly 19\% of the sentences in the corpus.} 
From these parses we ran our ILP and generated directionalized multiheaded dependency data to compare with the original conll data.

\begin{figure}[ht!]
  \centering
  \small
  \input{figure/conll_analysis_sentences.tex}
  \caption{\small Of the sentences used for this experiment section about \input{figure/conll_analysis_sentences_dropped}had at least one dropped word, and about \input{figure/conll_analysis_sentences_multiheaded}had at least one multiheaded word}
\end{figure}

@@

\TODO{more in the appendix ...}

@@

\section{Future Work}

\subsection{Slack Hierarchy}
We would like to introduce slack on the link grammar types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information. This slack would place a prior on rare or never-before seen link types to be assigned in the same way as other similar link types. We think that this will give the ILP better flexibility to handle noise and novel link types while still trying to learn the overall link grammar.

@@

\clearpage
\appendix
\section{English Link Types}

\TODO{A table here that lists each link grammar type together with its
  count, its allowed direction(s), its percentage of majority
  directions by token, and its most common corresponding CoNLL tags
  with percentages.  Readers could cut-and-paste this data to make a 
  cheap post-processor for LG parses.}

@@

\TODO{Show a bunch of English sentences with CoNLL vs.\@ link parses.
  Explain how they were selected and how the results are colorized etc.}

>>>>

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}  % allows better color names
\usepackage{todonotes}   % insert [disable] to disable all notes
% Note that these macros accept optional arguments such as 
% [size=\small,bordercolor=red].
\newcommand{\Note}[4][]{\todo[author=#2,color=#3,fancyline,#1]{#4}}
\newcommand{\noteJH}[2][]{\Note[#1]{MRG}{blue!40}{#2}}
\newcommand{\noteJE}[2][]{\Note[#1]{JE}{green!40}{#2}}   
\newcommand{\notewho}[3][]{\Note[#1]{#2}{orange!40}{#3}}  % extra arg with miscellaneous author
\newcommand{\NoteJH}[2][]{\noteJH[inline,#1]{#2}}
\newcommand{\NoteJE}[2][]{\noteJE[inline,#1]{#2}}
\newcommand{\Notewho}[3][]{\notewho[inline,#1]{#2}{#3}}  % extra arg with miscellaneous author

\title{Deriving Multi-Headed Projective Dependency Parses \\ from Link Grammar Parses}
% Oriented link grammar: Creating a multi-headed dependency corpus.

% \author{Juneki Hong and Jason Eisner\\
%   Department of Computer Science \\
%   Johns Hopkins University \\
%   Baltimore, MD 21218, USA \\ 
%   {\tt \{juneki,jason\}@cs.jhu.edu} \\
% }

@@

\begin{abstract}

% Copied from the google doc
Under multi-headed dependency grammar, a parse is a DAG rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, there are few multi-headed corpora at present, particularly for the projective case, so it is hard to investigate the benefit of such parsers or to work on making them faster or more accurate.  To fill this gap, we observe that link grammar produces {\em undirected} projective graphs.  
% OLD
% ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of !!! \noteJE{fill in the number} parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English, Russian, and more. \noteJE{what statistics does it use?} We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses. The resulting parses in English appear reasonably linguistically plausible, although they are not in general consistent with CoNLL-style parses of the same sentences; we discuss the differences.  We also report on the convergence of our ILP method to a stable annotation scheme as the corpus size increases.

\end{abstract}

\section{Introduction}

Dependency parsing is the task of mapping a sentence to a projective (not always projective?) directed acyclic tree. Link parsing in contrast produces a multiheaded planar graph with undirected edges, where every edge has a label describing the relationship between two constituents in a parse. In this paper we explore whether these relationships include dependencies. To determine the directional dependencies within the link edge labels we will use integer linear programming, encoding the problem in the Zimpl little language \cite{Koch2004}. It turns out that the link parses roughly only match half of the conll dependency corpus. However this is because \todo{}.

%Despite this, the link parses offer additional links in addition to dependencies...

Some recent work has considered multi-headed ...

 by expressing the dependency of a word on multiple parents.

\cite{mcdonald-pereira-2006,buchkromann-2006,sagae-tsujii-2008,gomezrodriguez-nivre-2013}

 (check this) except for Buch-Kromann's work on Danish \cite{kromann-2003}

@@

\section{Link Grammars}

Link grammars describe the links, or relationships between constituents in a parse \cite{SleatorTemperly91}.

We ran the link parser on the english bnews corpus sentences \footnote{We ignored the link parses that the link parser could not find suitable attachments and returned a disconnected graph. This happened for roughly NNN\% of the corpus.}

The link grammar on which the link-parser makes its linking decisions based on a set of handwritten dictionaries. Instead of going through these dictionaries, we learned the link grammar using an ILP. This approach allows us to analyze any link grammar dictionary other than English.

\section{Integer Linear Programming}

We formulated the ILP to assign directional dependencies to a link grammar parse such that it is a connected DAG, with all of the nodes reachable from the root.

The ILP chooses whether to set a link type to be assigned left, right, or both. For this, two label\_direction variables were made for each link type that appears in the parses, corresponding to whether all of the $l$ link edges were to be set to go left, right, or both. Because we assign a directionality to every link type, at least one of these variables must be set to true.

For example, setting the variables $$l_{LEFT} = TRUE, l_{RIGHT} = FALSE$$ would mean that the directional dependency of all links with label $l$ would be set to $LEFT$. If both variables were set to true, then the link type is allowed to go in either direction. Our ILP objective is to minimize the number of label\_direction variables set to true, while satisfying the other constraints.

\todo{cite model minimization papers}

\subsection{Slack Hierarchy}

We introduced slack on the directional dependency variable tokens such that link types were allowed to deviate from the majority up to (1\% ?) of the time before the ILP would assign both directional dependency variables to true. This addressed noise in the link parser's label assignments, while still allowing for the possibility that both directions could still be assigned.

\todo{} We also introduced slack on the link types such that link types with the same coarse grained label would try and align the same way as the majority in the group, where the preceding capital letters of the link type denote the coarse grained label, while the subscript letters denote further information. This slack places a prior on rare or never-before seen link types to be assigned in the same way as other similar link types.

This slack hierarchy gave the ILP the flexibility to handle noise and novel link types while still trying to learn the overall link grammar.

\begin{equation}
  count_{L, \textsc{left}} = \sum_{a \in \textsc{ARCS}_L} left_a
\end{equation}
\begin{equation}
  count_{L, \textsc{left}} = \textsc{M} \cdot \textsc{allowed}_{L,\textsc{left}} + \textsc{slack}_{L}
\end{equation}
Where $L \in \textsc{LABELS}$, and $\textsc{M}$ is the size of the corpus.
The objective is:
\begin{equation}
  \min \sum_{L \in \textsc{LABELS} \atop \textsc{DIR} \in \{left, right\}} (\textsc{allowed}_{L,\textsc{DIR}}) \atop + \textsc{COST}(L) * \sum_L \textsc{slack}_L
\end{equation}
\begin{equation}
  \textsc{COST}(L) = \frac{100}{(\textsc{count}_{L,\textsc{left}} + \textsc{count}_{L,\textsc{right}})}
\end{equation}

\begin{algorithm}
\caption{ILP Encoding}\label{ILP}
\begin{algorithmic}[1]
%\Procedure{MyProcedure}{}

\State $\text{min} \sum_{l \in labels} allowed_L(l) + allowed_R(l)$
\State $\text{subject to}$
\State $\text{Acyclicity}$
\State $\text{All nodes reachable from the root.}$

%\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}[ht!]
  \includegraphics[width=\linewidth, keepaspectratio=true]{figure/precision_recall.png}
  \caption{Precision and Recall of different runs on English}
\end{figure}

@@

\section{Other languages}

\section{Link Corpus}

@@

\todo{more in the appendix ...}

@@

@@

\appendix
\section{English Link Types}

\todo{A table here that lists each link grammar type together with its
  count, its allowed direction(s), its percentage of majority
  directions by token, and its most common corresponding CoNLL tags
  with percentages.  Readers could cut-and-paste this data to make a 
  cheap post-processor for LG parses.}

@@

\todo{Show a bunch of English sentences with CoNLL vs.\@ link parses.
  Explain how they were selected and how the results are colorized etc.}
